Finiteness in Cubical Type Theory

### 2020-01-25 paper.tex
finished a lot

When we say ``relation'' we are referring to the arrows in
figure~\ref{finite-classification}.
In section~\ref{relations} we will provide a function which inhabits each arrow.
Each unlabelled arrow is a weakening: i.e., every manifest Bishop finite set
is manifest enumerable (lemma~\ref{manifest-bishop-to-manifest-enum}).
The labelled arrows are strengthening proofs: i.e., every manifest enumerable set
\emph{with decidable equality} is manifest Bishop finite
(lemma~\ref{manifest-enum-to-manifest-bishop}).
Our most significant result here is the proof that a cardinal finite set with a
decidable total order is manifestly Bishop finite.

@@

\footnotetext{
  Actually, CuTT does have an identity type with similar semantics to the
  identity type in MLTT.
  We do not use this type anywhere in our work, however, so we will not consider
  it here.
}
\begin{romdefinition}[Homotopy Levels] \label{homotopy-types}
  Types in HoTT and CuTT are not necessarily sets, as they are in MLTT.
  This means that equalities are not necessarily unique.
  In fact, we have an entire hierarchy of homotopy levels, of which sets are
  level 2.
  \begin{alignat}{2}
    &\text{isContr}(A)    &&\coloneqq \Sigma(x : A) , \Pi(y : A) , (x \equiv y) \\
    &\text{isProp}(A)     &&\coloneqq \Pi(x, y : A) , (x \equiv y) \\
    &\text{isSet}(A)      &&\coloneqq \Pi(x, y : A) , \text{isProp}(x \equiv y) \\
    &\text{isGroupoid}(A) &&\coloneqq \Pi(x, y : A) , \text{isSet}(x \equiv y)
  \end{alignat}
  We can define the above types inductively like so:
  \begin{alignat}{2}
    &\text{isOfHLevel}(0 , A)   &&\coloneqq \text{isContr}(A) \\
    &\text{isOfHLevel}(n+1 , A) &&\coloneqq \Pi(x , y : A) , \text{isOfHLevel}(n , x \equiv y)
  \end{alignat}
  This type is defined similarly in \cite[definition 7.1.1]{hottbook}, although
  there the numbering starts at \(-2\).
\end{romdefinition}
\begin{romdefinition}[Fibres] \label{fibres}
  A fibre \cite[definition 4.2.4]{hottbook} is defined over some function \(f :
  A \rightarrow B\).
  \begin{equation}
    \text{fib}_f(y) = \Sigma(x : A) , (f x \equiv y)
  \end{equation}
\end{romdefinition}
\begin{romdefinition}[Equivalences] \label{equivalences}
  We will take contractible maps \cite[definition 4.4.1]{hottbook} as our
  ``default'' definition of equivalences.
  \begin{alignat}{2}
    &\text{isEquiv}(f) &&\coloneqq \Pi(y : B) , \text{isContr}(\text{fib}_f(y)) \label{is-equiv-def} \\
    &A \simeq B        &&\coloneqq \Sigma(f : A \rightarrow B) , \text{isEquiv}(f)
  \end{alignat}
\end{romdefinition}
\begin{romlemma}
  Univalence
  \begin{equation}
    (A \simeq B) \simeq (A \equiv B)
  \end{equation}
  This is the so-called univalence ``axiom'', which is a theorem in CuTT.
\end{romlemma}
\begin{romdefinition}[Decision]
  \begin{equation}
    \mathbf{Dec}(A) \coloneqq A \uplus \neg A
  \end{equation}
\end{romdefinition}
\begin{romdefinition}[Discrete]
  A discrete type is one with decidable equality.
  \begin{equation}
    \text{Discrete}(A) \coloneqq \Pi(x, y : A) , \mathbf{Dec}(x \equiv y)
  \end{equation}
  By Hedberg's theorem \cite{hedbergCoherenceTheoremMartinLof1998} any discrete
  type is a set.
\end{romdefinition}
\begin{romdefinition}[Higher Inductive Types] \label{HITs}
  Normal inductive types have \emph{point} constructors: constructors which
  construct values of the type.
  Higher Inductive Types (HITs) also have \emph{path} constructors: ways to
  construct paths in the type.
\end{romdefinition}
\begin{romdefinition}[Propositional Truncation] \label{prop-trunc}
  The type \(\lVert A \rVert\) on some type \(A\) is a propositionally truncated
  proof of \(A\) \cite[3.7]{hottbook}.
  In other words, it is a proof that some \(A\) exists, but it does not tell you
  \emph{which} \(A\).

@@

Where split enumerability was the enumeration form of a split surjection from
\(\mathbf{Fin}\), manifest Bishop finiteness is the enumeration form of an
\emph{equivalence} with \(\mathbf{Fin}\).
\begin{romlemma} \label{bishop-equiv}
  \begin{equation}
    \mathcal{B}(A) \simeq \Sigma {(n : \mathbb{N})} , \left( \mathbf{Fin}(n) \simeq A \right)
  \end{equation}
\end{romlemma}
\begin{proof}
  \begin{align*}
    \mathcal{B}(A)
    & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , x \inunique \mathit{xs}
    && \text{def.~\ref{bish-def} } (\mathcal{B})
    \\
    & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , \text{isContr} \left(\text{fib}_{\text{snd}(\mathit{xs})}(x)\right)
    && \text{def.~\ref{uniq-memb-def} } (\inunique)
    \\
    & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \text{isEquiv}(\text{snd}(\mathit{xs}))
    && \text{eqn.~\ref{is-equiv-def} (isEquiv)}
    \\
    & \simeq \Sigma (\mathit{xs} : \llbracket \mathbb{N} , \mathbf{Fin} \rrbracket (A)) , \text{isEquiv}(\text{snd}(\mathit{xs}))
    && \text{def.~\ref{List} } (\mathbf{List})
    \\
    & \simeq \Sigma (\mathit{xs} : \Sigma (n : \mathbb{N}) , \Pi (i : \mathbf{Fin}(n)) , A) , \text{isEquiv}(\text{snd}(\mathit{xs}))
    && \text{eqn.~\ref{container-interp} } (\llbracket \wc \rrbracket)
    \\
    & \simeq \Sigma (n : \mathbb{N}) , \Sigma (f : \mathbf{Fin}(n) \rightarrow A) , \text{isEquiv}(f)
    && \text{Reassociation of } \Sigma
    \\
    & \simeq \Sigma (n : \mathbb{N}) , ( \mathbf{Fin}(n) \simeq \; A )
    && \text{eqn.~\ref{surj-arrow-eqn} } (\twoheadrightarrow)
  \end{align*}
  \qed
\end{proof}
\subsection{Cardinal Finiteness}
Each finiteness predicate so far has contained an \emph{ordering} of the
underlying type.
For our purposes, this is too much information: it means that when constructing
the ``category of finite sets'' later on, instead of each type having one
canonical representative, it will have \(n!\), where \(n\) is the cardinality of
the type.

To remedy the problem, we will use propositional truncation
(def.~\ref{prop-trunc}).
\begin{romdefinition}[Cardinal Finiteness]
  \begin{equation}
    \mathcal{C}(A) \coloneqq \lVert \mathcal{B}(A) \rVert
  \end{equation}
\end{romdefinition}
At first glance, it might seem that we lose any useful properties we could
derive from \(\mathcal{B}\).
Luckily, this is not the case: by eliminator \ref{elim-prop-coh} of
def.~\ref{prop-trunc}, we need only show that the output is uniquely determined.

@@

The first thing we must define is a representation of subsets.
\begin{romdefinition}[Kuratowski Finite Subset]
  \(\mathcal{K}(A)\) is the type of Kuratowski-finite subsets of \(A\).
  \begin{equation}
    \begin{alignat}{2}
      \mathcal{K}(A) \coloneqq&
             \; []                &&: \mathcal{K}(A) ; \\
      \vert& \; \wc \dblcolon \wc &&: A \rightarrow \mathcal{K}(A) \rightarrow \mathcal{K}(A) ; \\
      \vert& \; \text{com}        &&: \Pi (x, y: A) , \Pi (\mathit{xs} : \mathcal{K}(A)) , x \dblcolon y \dblcolon \mathit{xs} \equiv y \dblcolon x \dblcolon \mathit{xs} ; \\
      \vert& \; \text{dup}        &&: \Pi (x : A) , \Pi (\mathit{xs} : \mathcal{K}(A)) , x \dblcolon x \dblcolon \mathit{xs} \equiv x \dblcolon \mathit{xs} ; \\
      \vert& \; \text{trunc}      &&: \Pi (\mathit{xs}, \mathit{ys}: \mathcal{K}(A)) , \Pi (p, q : \mathit{xs} \equiv \mathit{ys}) , p \equiv q ;
    \end{alignat}
  \end{equation}

  The first two constructors are point constructors, giving ways to create
  values of type \(\mathcal{K}(A)\).
  They are also recognisable as the two constructors for finite lists, a type
  which represents the free monoid.

@@

We also have the following two lemmas, proven in both
\cite{fruminFiniteSetsHomotopy2018} and our formalisation.
\begin{romlemma}
  \(\mathcal{K}^f\) is a mere proposition.
\end{romlemma}
\begin{romlemma}
  This circle \(S^1\) is Kuratowski finite.
\end{romlemma}
The second of these in particular tells us that the ``Kuratowski-finite types''
are not necessarily sets; it also tells us that we cannot derive decidable
equality from a proof of Kuratowski finiteness.
\section{Relations Between Each Finiteness Definition} \label{relations}
We now look at under what conditions a given finiteness predicate satisfies
another.
\subsection{Split Enumerability and Manifest Bishop Finiteness}
While manifest Bishop finiteness might seem stronger than split enumerability,
it turns out this is not the case.
Both types imply the other.
\begin{romlemma} \label{manifest-bishop-to-split-enum}
  Any manifest Bishop finite type is split enumerable.
\end{romlemma}

To derive \(\mathcal{B}\) from \(\mathcal{E}!\) takes significantly more work.
The ``unique membership'' condition in \(\mathcal{B}\) means that we are not
permitted duplicates in the support list.
The first step in the proof, then, is to filter those duplicates out from the
support list of the \(\mathcal{E}!\) proof: we can do this using the decidable
equality provided by \(\mathcal{E}!\) (lemma~\ref{split-enum-discrete}).
From there, all we need to show is that the membership proof carries over
appropriately.
\begin{romlemma} \label{split-enum-to-manifest-bishop}
  Any split enumerable set is manifest Bishop finite.
\end{romlemma}
\subsection{Manifest Bishop Finiteness and Manifest Enumerability}
\begin{romlemma} \label{manifest-bishop-to-manifest-enum}
  Any manifestly Bishop finite type is manifestly enumerable.
\end{romlemma}
\begin{romlemma} \label{manifest-enum-to-manifest-bishop}
  A manifestly enumerable type with decidable equality is manifest Bishop
  finite.
\end{romlemma}
\subsection{Manifest Bishop Finiteness and Cardinal Finiteness}
\begin{romlemma} \label{manifest-bishop-to-cardinal}
  Any manifest Bishop finite type is cardinal finite.
\end{romlemma}
\begin{romtheorem} \label{cardinal-to-manifest-bishop}
  Any cardinal finite type with a total order is Bishop finite.
\end{romtheorem}
\subsection{Cardinal Finiteness and Kuratowski Finiteness}
\begin{romlemma} \label{cardinal-kuratowski}
  \begin{equation}
    \mathcal{C}(A) \simeq \mathcal{K}(A) \times \text{Discrete}(A)
  \end{equation}
\end{romlemma}
\section{Topos} \label{topos}
Our main result in this section is the formal proof that decidable Kuratowski
finite types form a \(Pi\)-pretopos.
As we saw in theorem~\ref{kuratowski-cardinal}, decidable Kuratowski finite
types are equivalent to cardinal finite types, so we will work with the latter
from now on.
Our first task is to show that cardinal finite types are closed under several
operations.
\subsection{Closure}
For the first two closure proofs, we only consider manifest Bishop finiteness:
as it is the strongest of the finiteness predicates, we can derive the other
closure proofs from it.
\begin{romlemma}
  \(\bot\) is manifest Bishop finite.
\end{romlemma}
\begin{proof}
  From lemma~\ref{bishop-equiv} we can prove the goal by constructing a term of
  the type:
  \begin{equation}
    \Sigma(n : \mathbb{N}) , (\mathbf{Fin}(n) \simeq \bot)
  \end{equation}
  To do so, we take \(n\) to be 0, and by definition~\ref{Fin},
  \(\mathbf{Fin}(0) = \bot\).
  The equivalence is simply the identity equivalence.
\end{proof}
\begin{romlemma}
  \(\top\) is manifest Bishop finite.
\end{romlemma}
\begin{proof}
  Again, we need only provide a suitable cardinality (in this case 1), and an
  equivalence with \(\mathbf{Fin}(1)\) to prove our goal.
  \begin{align*}
    \top &\simeq \mathbf{Fin}(1) \\
          &\simeq \top \uplus \mathbf{Fin}(0) \\
          &\simeq \top \uplus \bot \\
          &\simeq \top
  \end{align*}
  \qed
\end{proof}
\begin{romlemma}
  \(\mathbf{Bool}\) is manifest Bishop finite.
\end{romlemma}

\begin{romlemma}
  Split enumerability is closed under \(\Sigma\),
\end{romlemma}
\begin{romlemma}
  Manifest bishop finiteness is closed over dependent functions
  (\(\prod\)-types).
  \begin{equation}
    \frac{
      \mathcal{B}(A) \; \; \; \Pi {(x : A)} , \mathcal{B}\left( U(x) \right)
    }{
      \mathcal{B}\left(\Pi {(x : A)} , U(x)\right)
    }
  \end{equation}
\end{romlemma}
\begin{proof}
  This proof is essentially the composition of two transport operations, made
  available to us via univalence.

  First, we will simplify things slightly by working only with split
  enumerability.
  As this is equal in strength to manifest Bishop finiteness, any closure proofs
  carry over.

  Secondly, we will replace \(A\) in all places with \(\mathbf{Fin}(n)\).
  Since we have already seen an equivalence between these two types, we are
  permitted to transport along these lines.
  This is the first transport operation.

  The bulk of the proof now is concerned with proving the following:
  \begin{equation}
    \left( \Pi {(x : \mathbf{Fin}(n))} , \mathcal{E}!(A(x)) \right) \rightarrow \mathcal{E}!\left( \Pi {(x : \mathbf{Fin}(n))} , A(x) \right)
  \end{equation}
  Our strategy to accomplish this will be to consider functions from
  \(\mathbf{Fin}(n)\) as \(n\)-tuples over some type family \(T : \mathbb{N}
  \rightarrow \text{Type}\).
  \begin{equation}
    \begin{aligned}
      \mathbf{Tuple}(T, 0)   &= \top \\
      \mathbf{Tuple}(T, n+1) &= T(0) \times \mathbf{Tuple}(T \circ \text{suc}, n)
    \end{aligned}
  \end{equation}
  This type is manifestly Bishop finite, as it is constructed only from products
  and the unit type.

  We then prove an isomorphism between this representation and \(\Pi\)-types.
  \begin{equation}
    \mathbf{Tuple}(T, n) \iff \Pi {(x : \mathbf{Fin}(n))} , T(x)
  \end{equation}
  This allows us to transport our proof of finiteness on tuples to one on
  functions from \(\mathbf{Fin}\) (our second transport operation), proving our
  goal.
\end{proof}
\subsection{The Category of Finite Sets}
HoTT and CuTT seem to be especially suitable settings for formalisations of
category theory.
The univalence axiom in particular allows us to treat categorical isomorphisms
as the important, useful objects that they are, upgrading them to full
equalities.

>>>>

When we say ``relation'' we are referring to the arrows in
figure~\ref{finite-classification}.
In section~\ref{relations} we will provide a function which inhabits each arrow.
Each unlabelled arrow is an unconditional implication: every manifest Bishop
finite set is cardinal finite (lemma~\ref{manifest-bishop-to-cardinal}), for
instance.
The labelled arrows are strengthening proofs: every manifest enumerable set
\emph{with decidable equality} is split enumerable
(lemma~\ref{manifest-enum-to-split-enum}).
Our most significant result here is the proof that a cardinal finite set with a
decidable total order is manifestly Bishop finite.

@@

\footnotetext{
  Actually, CuTT does have an identity type with similar semantics to the
  identity type in MLTT.
  We do not use this type anywhere in our work, however, so we will not consider
  it here.
}
\begin{romdefinition}[Homotopy Levels] \label{homotopy-types}
  Types in HoTT and CuTT are not necessarily sets, as they are in MLTT.
  This means that equalities are not necessarily unique.
  In fact, we have an entire hierarchy of homotopy levels, starting with
  contractions, of which sets are level 2.
  \begin{alignat}{2}
    &\text{isContr}(A)    &&\coloneqq \Sigma(x : A) , \Pi(y : A) , (x \equiv y) \\
    &\text{isProp}(A)     &&\coloneqq \Pi(x, y : A) , (x \equiv y) \\
    &\text{isSet}(A)      &&\coloneqq \Pi(x, y : A) , \text{isProp}(x \equiv y) \\
    &\text{isGroupoid}(A) &&\coloneqq \Pi(x, y : A) , \text{isSet}(x \equiv y)
  \end{alignat}
  We can define the above types inductively like so:
  \begin{alignat}{2}
    &\text{isOfHLevel}(0 , A)   &&\coloneqq \text{isContr}(A) \\
    &\text{isOfHLevel}(n+1 , A) &&\coloneqq \Pi(x , y : A) , \text{isOfHLevel}(n , x \equiv y)
  \end{alignat}
  This type is defined similarly in \cite[definition 7.1.1]{hottbook}, although
  there the numbering starts at \(-2\).
\end{romdefinition}
\begin{romdefinition}[Fibres] \label{fibres}
  A fibre \cite[definition 4.2.4]{hottbook} is defined over some function \(f :
  A \rightarrow B\).
  \begin{equation}
    \text{fib}_f(y) = \Sigma(x : A) , (f (x) \equiv y)
  \end{equation}
\end{romdefinition}
\begin{romdefinition}[Equivalences] \label{equivalences}
  We will take contractible maps \cite[definition 4.4.1]{hottbook} as our
  ``default'' definition of equivalences.
  \begin{alignat}{2}
    &\text{isEquiv}(f) &&\coloneqq \Pi(y : B) , \text{isContr}(\text{fib}_f(y)) \label{is-equiv-def} \\
    &A \simeq B        &&\coloneqq \Sigma(f : A \rightarrow B) , \text{isEquiv}(f)
  \end{alignat}
\end{romdefinition}
\begin{romlemma}
  Univalence
  \begin{equation}
    (A \simeq B) \simeq (A \equiv B)
  \end{equation}
  This is the so-called univalence ``axiom'', which is a theorem in CuTT.
\end{romlemma}
\begin{romdefinition}[Decision]
  \begin{equation}
    \mathbf{Dec}(A) \coloneqq A \uplus \neg A
  \end{equation}
\end{romdefinition}
\begin{romdefinition}[Discrete]
  A discrete type is one with decidable equality.
  \begin{equation}
    \text{Discrete}(A) \coloneqq \Pi(x, y : A) , \mathbf{Dec}(x \equiv y)
  \end{equation}
  By Hedberg's theorem \cite{hedbergCoherenceTheoremMartinLof1998} any discrete
  type is a set.
\end{romdefinition}
\begin{romdefinition}[Higher Inductive Types] \label{HITs}
  Normal inductive types have \emph{point} constructors: constructors which
  construct values of the type.
  Higher Inductive Types (HITs) also have \emph{path} constructors: ways to
  construct paths in the type.
\end{romdefinition}
\begin{romdefinition}[Propositional Truncation] \label{prop-trunc}
  The type \(\lVert A \rVert\) on some type \(A\) is a propositionally truncated
  proof of \(A\) \cite[3.7]{hottbook}.
  In other words, it is a proof that some \(A\) exists, but it does not tell you
  \emph{which} \(A\).

@@

Where split enumerability was the enumeration form of a split surjection from
\(\mathbf{Fin}\), manifest Bishop finiteness is the enumeration form of an
\emph{equivalence} with \(\mathbf{Fin}\).
\begin{romlemma} \label{bishop-equiv}
  \begin{equation}
    \mathcal{B}(A) \simeq \Sigma {(n : \mathbb{N})} , \left( \mathbf{Fin}(n) \simeq A \right)
  \end{equation}
\end{romlemma}
\begin{proof}
  \begin{align*}
    \mathcal{B}(A)
    & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , x \inunique \mathit{xs}
    && \text{def.~\ref{bish-def} } (\mathcal{B})
    \\
    & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , \text{isContr} \left(\text{fib}_{\text{snd}(\mathit{xs})}(x)\right)
    && \text{def.~\ref{uniq-memb-def} } (\inunique)
    \\
    & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \text{isEquiv}(\text{snd}(\mathit{xs}))
    && \text{eqn.~\ref{is-equiv-def} (isEquiv)}
    \\
    & \simeq \Sigma (\mathit{xs} : \llbracket \mathbb{N} , \mathbf{Fin} \rrbracket (A)) , \text{isEquiv}(\text{snd}(\mathit{xs}))
    && \text{def.~\ref{List} } (\mathbf{List})
    \\
    & \simeq \Sigma (\mathit{xs} : \Sigma (n : \mathbb{N}) , \Pi (i : \mathbf{Fin}(n)) , A) , \text{isEquiv}(\text{snd}(\mathit{xs}))
    && \text{eqn.~\ref{container-interp} } (\llbracket \wc \rrbracket)
    \\
    & \simeq \Sigma (n : \mathbb{N}) , \Sigma (f : \mathbf{Fin}(n) \rightarrow A) , \text{isEquiv}(f)
    && \text{Reassociation of } \Sigma
    \\
    & \simeq \Sigma (n : \mathbb{N}) , ( \mathbf{Fin}(n) \simeq \; A )
    && \text{eqn.~\ref{surj-arrow-eqn} } (\twoheadrightarrow)
  \end{align*}
  \qed
\end{proof}
\subsection{Cardinal Finiteness}
Each finiteness predicate so far has contained an \emph{ordering} of the
underlying type.
For our purposes, this is too much information: it means that when constructing
the ``category of finite sets'' later on, instead of each type having one
canonical representative, it will have \(n!\), where \(n\) is the cardinality of
the type\footnotemark.

\footnotetext{
  We actually do get a category (a groupoid, even) from manifest Bishop
  finiteness \cite{yorgeyCombinatorialSpeciesLabelled2014}: it's the groupoid of
  finite sets equipped with a linear order, whose morphisms are order-preserving
  bijections.
  We do not explore this particular construction in any detail.
}

To remedy the problem, we will use propositional truncation
(def.~\ref{prop-trunc}).
\begin{romdefinition}[Cardinal Finiteness]
  \begin{equation}
    \mathcal{C}(A) \coloneqq \lVert \mathcal{B}(A) \rVert \simeq \lVert \Sigma(n : \mathbb{N}) , (\mathbf{Fin}(n) \simeq A) \rVert
  \end{equation}
\end{romdefinition}
At first glance, it might seem that we lose any useful properties we could
derive from \(\mathcal{B}\).
Luckily, this is not the case: by eliminator \ref{elim-prop-coh} of
def.~\ref{prop-trunc}, we need only show that the output is uniquely determined.

@@

The first thing we must define is a representation of subsets.
\begin{romdefinition}[Kuratowski Finite Subset]
  \(\mathcal{K}(A)\) is the type of Kuratowski-finite subsets of \(A\).
  \begin{equation}
    \begin{alignat}{2}
      \mathcal{K}(A) \coloneqq&
             \; []                &&: \mathcal{K}(A) ; \\
      \vert& \; \wc \dblcolon \wc &&: A \rightarrow \mathcal{K}(A) \rightarrow \mathcal{K}(A) ; \\
      \vert& \; \text{com}        &&: \Pi (x, y: A) , \Pi (\mathit{xs} : \mathcal{K}(A)) , x \dblcolon y \dblcolon \mathit{xs} \equiv y \dblcolon x \dblcolon \mathit{xs} ; \\
      \vert& \; \text{dup}        &&: \Pi (x : A) , \Pi (\mathit{xs} : \mathcal{K}(A)) , x \dblcolon x \dblcolon \mathit{xs} \equiv x \dblcolon \mathit{xs} ; \\
      \vert& \; \text{trunc}      &&: \Pi (\mathit{xs}, \mathit{ys}: \mathcal{K}(A)) , \Pi (p, q : \mathit{xs} \equiv \mathit{ys}) , p \equiv q ;
    \end{alignat}
  \end{equation}
  We define it as a HIT (definition~\ref{HITs}).
  The first two constructors are point constructors, giving ways to create
  values of type \(\mathcal{K}(A)\).
  They are also recognisable as the two constructors for finite lists, a type
  which represents the free monoid.

@@

We also have the following two lemmas, proven in both
\cite{fruminFiniteSetsHomotopy2018} and our formalisation.
\begin{romlemma}
  \(\mathcal{K}^f\) is a mere proposition.
\end{romlemma}
\begin{romlemma}
  This circle \(S^1\) is Kuratowski finite.
\end{romlemma}
The second of these in particular tells us that the ``Kuratowski-finite types''
are not necessarily sets; it also tells us that we cannot derive decidable
equality from a proof of Kuratowski finiteness.
\section{Relations Between Each Finiteness Definition} \label{relations}
We will now look at the arrows in figure~\ref{finite-classification}.
\subsection{Split Enumerability and Manifest Bishop Finiteness}
While manifest Bishop finiteness might seem stronger than split enumerability,
it turns out this is not the case.
Both types imply the other.
\begin{romlemma} \label{manifest-bishop-to-split-enum}
  Any manifest Bishop finite type is split enumerable.
\end{romlemma}
\begin{proof}
  To construct a proof of split enumerability from one of manifest Bishop
  finiteness, it suffices to convert a proof of \(x \inunique \mathit{xs}\) to
  one of \(x \in \mathit{xs}\), for all \(x\) and \(\mathit{xs}\).
  Since \(\inunique\) is defined as a contraction of \(\in\), such a conversion
  is simply the \(\text{fst}\) function.
  \qed
\end{proof}

To derive \(\mathcal{B}\) from \(\mathcal{E}!\) takes significantly more work.
The ``unique membership'' condition in \(\mathcal{B}\) means that we are not
permitted duplicates in the support list.
The first step in the proof, then, is to filter those duplicates out from the
support list of the \(\mathcal{E}!\) proof: we can do this using the decidable
equality provided by \(\mathcal{E}!\) (lemma~\ref{split-enum-discrete}).
From there, all we need to show is that the membership proof carries over
appropriately.
\begin{romlemma} \label{split-enum-to-manifest-bishop}
  Any split enumerable set is manifest Bishop finite.
\end{romlemma}
\subsection{Split Enumerability and Manifest Enumerability}
\begin{romlemma} \label{split-enum-to-manifest-enum}
  Any split enumerable type is manifestly enumerable.
\end{romlemma}
\begin{romlemma} \label{dec-trunc-refute}
  For any type \(A\), we can extract a value of type \(A\) from a decision over
  \(A\) in combination with a propositionally-truncated value of type \(\lVert A \rVert\).
  \begin{equation}
    \frac{
      \mathbf{Dec}(A) \; \; \; \lVert A \rVert
    }{
      A
    }
  \end{equation}
\end{romlemma}
\begin{proof}
  Let \(d\) be a value of type \(\mathbf{Dec}(A)\), and \(p\) be a value of type
  \(\lVert A \rVert\).
  We proceed by case analysis on \(d\).
  In the case where \(d\) is inl, we have a value of type \(A\), and our goal is
  satisfied.
  In the case of inr, we have a value of type \(\neg A\).
  Using eliminator \ref{elim-prop-prop} from definition~\ref{prop-trunc}, and
  the fact that \(\bot\) is a proposition, we can run the \(\neg A\) proof on
  \(p\), getting a contradiction.
  \qed
\end{proof}
\begin{romlemma} \label{manifest-enum-to-split-enum}
  A manifestly enumerable type with decidable equality is split enumerable.
\end{romlemma}
\begin{proof}
  Similarly to the proof of lemma~\ref{manifest-bishop-to-split-enum}, our
  obligation for this proof is to provide a conversion function between
  membership proofs.
  In this case, we need to show:
  \begin{equation}
    \Pi(x : A) , \Pi(\mathit{xs} : \mathbf{List}(A)) , \lVert x \in \mathit{xs} \rVert \rightarrow x \in \mathit{xs}
  \end{equation}
  This conversion function can be constructed from the two following functions,
  provided in the formalisation.
  First, for any type with decidable equality, decidable membership in a finite
  list is consequently decidable:
  \begin{equation}
    \text{Discrete}(A) \rightarrow \Pi(x : A) , \Pi(\mathit{xs} : \mathbf{List}(A)) , \mathbf{Dec}(x \in \mathit{xs})
  \end{equation}
  Then, we use lemma~\ref{dec-trunc-refute}, allowing us to derive \(x \in
  \mathit{xs}\) from \(\lVert x \in \mathit{xs} \rVert\).
  \qed
\end{proof}
\subsection{Manifest Bishop Finiteness and Cardinal Finiteness}
\begin{romlemma} \label{manifest-bishop-to-cardinal}
  Any manifest Bishop finite type is cardinal finite.
\end{romlemma}
\begin{romtheorem} \label{cardinal-to-manifest-bishop}
  Any cardinal finite type with a total order is Bishop finite.
\end{romtheorem}
The proof for this particular theorem is quite involved in the formalisation, so
we only give its sketch here.
First, note that we actually convert to manifest enumerability first: this can
be converted to split enumerability with decidable equality, which is provided
by cardinal finiteness.

Next, we define permutations.
\begin{romdefinition}[List Permutations]
  Two lists are permutations of each other if their membership proofs are all
  equivalent\footnotemark \cite{danielssonBagEquivalenceProofRelevant2012}.
  \begin{equation}
    \mathit{xs} \leftrightsquigarrow \mathit{ys} = \Pi {(x : A)} , x \in \mathit{xs} \simeq x \in \mathit{ys}
  \end{equation}
\end{romdefinition}

\footnotetext{
  The definition in \cite{danielssonBagEquivalenceProofRelevant2012} and our
  formalisation is slightly different: we say permutations are lists with
  \emph{isomorphic} membership proofs.
  The distinction, as it happens, does not affect our work here.
}

Next, we define a sort function which relies on the provided total order.
We further prove the following fact about this sort function:
\begin{equation}
  \Pi(\mathit{xs}, \mathit{ys} : \mathbf{List}(A)) , \mathit{xs} \leftrightsquigarrow \mathit{ys} \rightarrow \text{sort}(\mathit{xs}) \equiv \text{sort}(\mathit{ys})
\end{equation}

Next, notice that the support lists of any two proofs of manifest Bishop
finiteness must be permutations of each other.
This will allow us to sort the support list of a proof of cardinal finiteness in
a coherently constant (definition~\ref{prop-trunc},
eliminator~\ref{elim-prop-coh}) way, pulling the support list out from the
truncation.
The cover proof emerges naturally from the definition of the permutation.
\subsection{Cardinal Finiteness and Kuratowski Finiteness}
\begin{romlemma} \label{cardinal-kuratowski}
  \begin{equation}
    \mathcal{C}(A) \simeq \mathcal{K}(A) \times \text{Discrete}(A)
  \end{equation}
\end{romlemma}
This proof is constructed by providing a pair of functions: one from
\(\mathcal{C}(A)\) to \(\mathcal{K}(A) \times \text{Discrete}(A)\), and one the
other way.
This pair implies an equivalence, because both source and target are
propositions.
The actual functions themselves are proven in our formalisation, as well as in
\cite{fruminFiniteSetsHomotopy2018}.
\section{Topos} \label{topos}
In this section we will prove that decidable Kuratowski finite types form a
\(\Pi\)-pretopos.
Along the way we will provide closure proofs for a number of the other
finiteness predicates.
As we saw in theorem~\ref{cardinal-kuratowski}, decidable Kuratowski finite
types are equivalent to cardinal finite types, so we will work with the latter
from now on.
Our first task is to show that cardinal finite types are closed under several
operations.
\subsection{Closure}
For the first two closure proofs, we only consider manifest Bishop finiteness:
as it is the strongest of the finiteness predicates, we can derive the other
closure proofs from it.
\begin{romlemma}
  \(\bot\), \(\top\), and \(\mathbf{Bool}\) are manifest Bishop finite.
\end{romlemma}
\begin{romlemma}
  Split enumerability is closed under \(\Sigma\)
\end{romlemma}
To show closure under \(\Sigma\), we take the Cartesian product of the support
lists.

\begin{romlemma}
  Manifest bishop finiteness is closed over dependent functions
  (\(\prod\)-types).
  \begin{equation}
    \frac{
      \mathcal{B}(A) \; \; \; \Pi {(x : A)} , \mathcal{B}\left( U(x) \right)
    }{
      \mathcal{B}\left(\Pi {(x : A)} , U(x)\right)
    }
  \end{equation}
\end{romlemma}
\begin{proof}
  Firstly, we will simplify things slightly by working only with split
  enumerability.
  As this is equal in strength to manifest Bishop finiteness, any closure proofs
  carry over.
  Secondly, we will replace \(A\) in all places with \(\mathbf{Fin}(n)\), using
  the equivalence provided by lemma~\ref{bishop-equiv}.

  The remainder of the proof now is concerned with proving the following:
  \begin{equation} \label{given-transp-sublemma}
    \left( \Pi {(x : \mathbf{Fin}(n))} , \mathcal{E}!(U(x)) \right) \rightarrow \mathcal{E}!\left( \Pi {(x : \mathbf{Fin}(n))} , U(x) \right)
  \end{equation}

  To prove this goal, consider the following type of \(n\)-tuples.
  \begin{equation}
    \begin{alignat}{3}
      & \mathbf{Tuple}(T, 0)   &&\coloneqq \top \\
      & \mathbf{Tuple}(T, n+1) &&\coloneqq T(0) \times \mathbf{Tuple}(T \circ \text{suc}, n)
    \end{alignat}
  \end{equation}
  We have the following equivalence, proven in our formalisation.
  \begin{equation} \label{fin-equiv-sublemma}
    \mathbf{Tuple}(T, n) \simeq \Pi {(x : \mathbf{Fin}(n))} , T(x)
  \end{equation}

  By the given proof of finiteness in equation~\ref{given-transp-sublemma}, we
  know that \(\mathbf{Tuple}(T , n)\) is split enumerable, since it is made of
  products of the points of \(U\).
  We can then transport along the equivalence in
  equation~\ref{fin-equiv-sublemma}, proving our goal.
\end{proof}

The dependent cases of each of these proofs do not transfer simply to proofs on
\(\mathcal{C}\).
In order to transfer them over, we need to prove a variant of the axiom of
choice on finite sets.
\begin{romlemma}
  \begin{equation}
    \mathcal{C}(A) \rightarrow (\Pi(x : A) , \lVert U(x) \rVert) \rightarrow \lVert \Pi(x : A) , U(x) \rVert
  \end{equation}
\end{romlemma}
This is proven in our formalisation.
\subsection{The Category of Finite Sets}
HoTT and CuTT seem to be especially suitable settings for formalisations of
category theory.
The univalence axiom in particular allows us to treat categorical isomorphisms
as equalities, saving us from the dreaded ``setoid hell''.
